Empirical normative theories

### 2020-08-02 simple.tex
Better notations and removed q(i) parameterisation of a theory

\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[version=last, pagesize, twoside=off, bibliography=totoc, DIV=calc, fontsize=12pt, a4paper, french, english]{scrartcl}
\input{preamble/packages}
\input{preamble/math_basics}
\input{preamble/math_mine}
\input{preamble/redac}
%\input{preamble/draw}
%\input{preamble/acronyms}
\addtokomafont{labelinglabel}{\sffamily\bfseries}
\renewcommand{\phi}{\varphi}%TODO
\DeclareMathAlphabet{\mathup}{OT1}{\familydefault}{m}{n}

@@

Given $i \in I, \phi \in \Phi, \hist \in \allhist$, say that $\hist$ leads strongly to $\phi$ iff $\hist \ileadsto \phi \land \hist \nileadsto ¬\phi$.
Define $\hist \ileadstost \phi$ iff $[\hist \text{ leads strongly to } \phi] \land [\phi = \propsure ⇒ \hist \text{ leads strongly to } \propposs]$.
Given $\hist \in \allhist$, let $\histend$ denote the set containing the last two elements of the sequence $\hist$ if the sequence has at least two elements; the singleton set containing the last element of the sequence if it has exactly one element; and the empty set if $\hist = \emptyset$.
Given $i \in I, \phi \in \Phi, \ar \in \allargs$, define $\ar \ileadstost \phi$ iff $\forall \hist \in \allhist \suchthat \ar \in \histend: \hist \ileadstost \phi$.

@@

\subsection{Non-completeness}
\begin{definition}[Theory completeness]
	$\forall \phi \in \Phi: [\exists \ar \in \allargs \suchthat \ar \ileadsto \phi] \lor [\exists \ar \in \allargs \suchthat \ar \ileadsto ¬\phi]$.
\end{definition}
\begin{definition}[Unrestricted interpretability]
	$\forall \phi \in \Phi: \phi \in \iPhi \oplus ¬\phi \in \iPhi$.
\end{definition}
\begin{definition}[Deliberation completeness]
	$\forall \phi \in \Phi: \phi \in \iPhi \lor ¬\phi \in \iPhi$.
\end{definition}
The following proposition directly follows from the definitions, together with \cref{th:protcoh}.
\begin{proposition}
	1) Under Normative adequacy, Theory completeness implies Deliberation completeness. 2) Unrestricted interpretability implies Deliberation completeness.
\end{proposition}
Theories as defined here do not mandate Theory completeness. This is good for two reasons. First, even the weaker Deliberation completeness can be considered too strong. Second, it is dubious that, for non trivial decision problems, theories can be found for which both Normative adequacy is normatively compelling and Theory completeness holds.

\section{Additive single-answer empirical theories}
\subsection{Definition}
\commentOC{I leave the definition as is, but for the moment I only consider ${\gleadsto}$, a relation that does not depend on $i$; and I do not use $q(i)$.}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, \alldargs, \dbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq (\allargs × \Phi) \cap (\alldargs × \Phi)$, 
${\gbeats} \subseteq \allargs × \allargs$ and is irreflexive,
${\dbeats} \subseteq \alldargs × \allargs$ and is injective and its range includes $\gbeatsinv(\allargs)$ (thus $\forall \ar_1 \suchthat [\exists \ar_0 \suchthat \ar_1 \gbeats \ar_0]$: $\dbeatsinv(\ar_1)$ is a singleton). 
Define $\df: \allargs → \powersetz{\allargs}$ as $\df = \dbeatsinv \circ \gbeatsinv$.
Given $\ar \in \allargs$, $\df(\ar)$ denotes the \emph{fragments defending $\ar$}. 
In other words, given $\ard \in \dargs$ and $\ar_0 \in \allargs$, $\ar_2 \in \df(\ar_0)$ iff $\exists \ar_1 \in \allargs \suchthat \ar_2 \dbeats \ar_1 \gbeats \ar_0$.
A \emph{coalition defending $\ar$} is any element of $\powersetz{\df(\ar)}$.
The symbol $+$ in an empirical theory stands for a function $+: \bigcup_{\ar \in \allargs}(\set{s} × \powersetz{\df(\ar)}) → \allargs$.
Given any $\ar \in \allargs$ and $\emptyset ≠ \dargs \subseteq \df(\ar)$, $\ar + \dargs$ means $+(\ar, \dargs)$. 
Given any $\ar \in \allargs$, define $\dfp(\ar)$, the \emph{defenders of $\ar$}, as $\dfp(\ar) = \set{\ar + \dargs \suchthat \emptyset ≠ \dargs \subseteq \df(\ar)}$.

@@

Given $i \in I, \phi \in \Phi, \ar \in \allargs$ and $\args \subseteq \gbeatsinv(\ar)$, define $\args \ibeats \ar$ iff $\exists \hist \in \allhist \suchthat \args = \gbeatsinv(\ar) \cap \hist \land \ar \in \histend \land \hist \nileadstost \phi$.

@@

Given $\phi \in \Phi$, $\hist \in \allhist$ and $\ar_\gamma \in \gargs$, say that $(\hist, \ar_\gamma)$ challenges $(\gamma, \phi)$ iff $\hist \nileadstost \phi$ and $\ar_\gamma \in \histend$.
Given $\phi \in \Phi, \ar_\gamma \in \gargs, \ar_\delta \in \gargsdelta$, say that $(\hist, \ar_\gamma, \ar_\delta)$ is a $(\gamma, \phi$ or $\delta, ¬\phi)$-challenge iff $(\hist, \ar_\gamma)$ challenges $(\gamma, \phi)$ or $(\hist, \ar_\delta)$ challenges $(\delta, ¬\phi)$. \commentOC{This should be called a $\phi$-or-$¬\phi$-challenge.}

@@

\begin{lemma}
	\label{th:nextchal}
	Given $(\hist, \ar_\gamma)$ challenging $(\gamma, \phi)$, with $\ar_\gamma \in \gargs \cap \dfpclosure$, and $\ar_\delta \in \gargsdelta$: either $(i, \phi, \hist, \ar_\gamma)$ falsifies $\gamma$, or there exists a $(\gamma, \phi$ or $\delta, ¬\phi)$-challenge $(\histp, \ar'_\gamma, \ar_\delta)$, with $\ar'_\gamma \in \gargs \cap (\dfp)^{k + 1}(\gleadstoinv(\phi))$.
\end{lemma}
\begin{proof}
	If $\ar_\gamma$ does not $i$-defend $\phi$, then $(i, \phi, \hist, \ar_\gamma)$ falsifies $\gamma$, which ends the proof. 
	Thus, assume that $\ar_\gamma$ $i$-defends $\phi$.

@@

\begin{theorem}
	Given two incompatible empirical theories for a given normative theory, $\alpha = (q, (\gleadsto[q(i)][\alpha])_{q(i) \in q(I)}, \gbeats[\alpha], \alldargs, \dbeats[\alpha], +_\alpha)$ and $\beta = (q', (\gleadsto[q(i)][\beta])_{q(i) \in q(I)}, \gbeats[\beta], \alldargs, \dbeats[\beta], +')$ and given any $i \in I$, there exists a test procedure with $i$ that falsifies one of these theories.
\end{theorem}
\begin{proof}
	Pick any $\phi \in \Phi$ about which $\alpha$ and $\beta$ disagree.

@@

	Pick any $\ar_{-1} \in \allargs \suchthat \ar_{-1} \gleadsto[][\beta] ¬\phi$. Observe that $\ar_{-1} \in \gargsbeta \cap (\dfp[\beta])^0(\gleadstoinv[\beta](¬\phi))$. Test whether $(\ar_{-1}) \nileadstost \phi$. If so, $(i, \phi, (\ar_{-1}), \ar_{-1})$ falsifies $\beta$, which ends the procedure. 
	Thus, assume that $(\ar_{-1}) \ileadstost \phi$.

	Pick any $\ar_0 \in \allargs \suchthat \ar_0 \gleadsto[][\alpha] \phi$. Observe that $\ar_0 \in \gargsalpha \cap (\dfp[\alpha])^0(\gleadstoinv[\alpha](\phi))$.  Define $\hist = (\ar_{-1}, \ar_0)$ and test whether $\hist \nileadstost \phi$. By \cref{th:endchal}, $(\hist, \ar_0, \ar_{-1})$ is a $(\alpha, \phi$ or $\beta, ¬\phi)$-challenge, thus, $(\hist, \ar_0, \ar_{-1})$ is a challenge of complexity $(0, 0)$.

@@

Define $\ileadsto$ as follows. Given $\hist = (…, \ar)$, $\hist \ileadstost \phi$ iff, in a situation of $i$ having been exposed to the prefix …, saying: “Considering the argument $\ar$, would you please indicate your maximum selling price for $P$, and for $D$” leads $i$ to act accordig to $\phi$. Note that $\hist \ileadsto \phi$ iff $\hist \nileadsto ¬\phi$.

>>>>

\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[version=last, pagesize, twoside=off, bibliography=totoc, DIV=calc, fontsize=12pt, a4paper, french, english]{scrartcl}
\input{preamble/packages}
\input{preamble/math_basics}
\input{preamble/math_mine}
\input{preamble/redac}
\usepackage[normalem]{ulem}
%\input{preamble/draw}
%\input{preamble/acronyms}
\addtokomafont{labelinglabel}{\sffamily\bfseries}
\renewcommand{\phi}{\varphi}%TODO
\DeclareMathAlphabet{\mathup}{OT1}{\familydefault}{m}{n}

@@

Given $i \in I, \phi \in \Phi, \hist \in \allhist$, say that $\hist$ leads strongly to $\phi$ iff $\hist \ileadsto \phi \land \hist \nileadsto ¬\phi$.
Define $\hist \ileadstost \phi$ iff $[\hist \text{ leads strongly to } \phi] \land [\phi = \propsure ⇒ \hist \text{ leads strongly to } \propposs]$.
(The visual proximity of the symbols $\ileadsto$ and $\ileadstost$ creates a small risk of confusion, but I feel it is justified by the fact that $\ileadsto$ is not used in the rest of this article, except in the proof of \cref{th:protcoh}.)
Given $\hist \in \allhist$, let $\histend$ denote the set containing the last two elements of the sequence $\hist$ if the sequence has at least two elements; the singleton set containing the last element of the sequence if it has exactly one element; and the empty set if $\hist = \emptyset$.
Given $i \in I, \phi \in \Phi, \ar \in \allargs$, define $\ar \ileadstost \phi$ iff $\forall \hist \in \allhist \suchthat \ar \in \histend: \hist \ileadstost \phi$.

@@

\subsection{Non-completeness}
\begin{definition}[Theory completeness]
	$\forall \phi \in \Phi: [\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] \lor [\exists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi]$.
\end{definition}
\begin{definition}[Unrestricted interpretability]
	$\forall \phi \in \Phi: \phi \in \iPhi \oplus ¬\phi \in \iPhi$.
\end{definition}
\begin{definition}[Deliberation completeness]
	$\forall \phi \in \Phi: \phi \in \iPhi \lor ¬\phi \in \iPhi$.
\end{definition}
The following proposition directly follows from the definitions, together with \cref{th:protcoh}.
\begin{proposition}
	1) Under Normative adequacy, Theory completeness implies Deliberation completeness. 2) Unrestricted interpretability implies Deliberation completeness.
\end{proposition}
Theories as defined here do not mandate Theory completeness. This is good for two reasons. First, even the weaker Deliberation completeness can be considered too strong. Second, it is dubious that, for non trivial decision problems, theories can be found for which both Normative adequacy is normatively compelling and Theory completeness holds.

\section{Additive single-answer empirical theories}
\subsection{Definition}
An empirical theory for a normative theory is a tuple $(\gleadsto, \gbeats, \alldargs, \dbeats, +)$,
with ${\gleadsto} \subseteq \allargs × \Phi$, 
${\gbeats} \subseteq \allargs × \allargs$ and is irreflexive,
${\dbeats} \subseteq \alldargs × \allargs$ and is injective and its range includes $\gbeatsinv(\allargs)$ (thus $\forall \ar_1 \suchthat [\exists \ar_0 \suchthat \ar_1 \gbeats \ar_0]$: $\dbeatsinv(\ar_1)$ is a singleton). 
Define $\df: \allargs → \powersetz{\allargs}$ as $\df = \dbeatsinv \circ \gbeatsinv$.
Given $\ar \in \allargs$, $\df(\ar)$ denotes the \emph{fragments defending $\ar$}. 
In other words, given $\ard \in \dargs$ and $\ar_0 \in \allargs$, $\ar_2 \in \df(\ar_0)$ iff $\exists \ar_1 \in \allargs \suchthat \ar_2 \dbeats \ar_1 \gbeats \ar_0$.
A \emph{coalition defending $\ar$} is any element of $\powersetz{\df(\ar)}$.
The symbol $+$ in an empirical theory stands for a function $+: \bigcup_{\ar \in \allargs}(\set{s} × \powersetz{\df(\ar)}) → \allargs$.
Given any $\ar \in \allargs$ and $\emptyset ≠ \dargs \subseteq \df(\ar)$, $\ar + \dargs$ means $+(\ar, \dargs)$. 
Given any $\ar \in \allargs$, define $\dfp(\ar)$, the \emph{defenders of $\ar$}, as $\dfp(\ar) = \set{\ar + \dargs \suchthat \emptyset ≠ \dargs \subseteq \df(\ar)}$.

@@

Given $i \in I, \phi \in \Phi, \ar \in \allargs$ and $\args \subseteq \gbeatsinv(\ar)$, define $\args \ibeats \ar$ iff $\exists \hist \in \allhist \suchthat \args = \gbeatsinv(\ar) \cap \hist \land \ar \in \histend \land \hist \nileadstost \phi$. (The striked through $\phi$ in superscript is there to suggest that, intuitively, $\args$ argue against $\phi$, from the point of view of $i$.)

@@

Given $\phi \in \Phi$, $\hist \in \allhist$ and $\ar_\gamma \in \gargs$, say that $(\hist, \ar_\gamma)$ challenges $(\gamma, \phi)$ iff $\hist \nileadstost \phi$ and $\ar_\gamma \in \histend$.
Given $\phi \in \Phi, \ar_\gamma \in \gargs, \ar_\delta \in \gargsdelta$, say that $(\hist, \ar_\gamma, \ar_\delta)$ is a $(\gamma, \phi$ or $\delta, ¬\phi)$-challenge iff $(\hist, \ar_\gamma)$ challenges $(\gamma, \phi)$ or $(\hist, \ar_\delta)$ challenges $(\delta, ¬\phi)$. \commentOC{This should be called a $\phi$/$¬\phi$-challenge.}

@@

\begin{lemma}
	\label{th:nextchal}
	Given $(\hist, \ar_\gamma)$ challenging $(\gamma, \phi)$, with $\ar_\gamma \in \gargs \cap (\dfp)^k(\gleadstoinv(\phi))$, and $\ar_\delta \in \gargsdelta$: either $(i, \phi, \hist, \ar_\gamma)$ falsifies $\gamma$, or there exists a $(\gamma, \phi$ or $\delta, ¬\phi)$-challenge $(\histp, \ar'_\gamma, \ar_\delta)$, with $\ar'_\gamma \in \gargs \cap (\dfp)^{k + 1}(\gleadstoinv(\phi))$.
\end{lemma}
\begin{proof}
	If $\ar_\gamma$ does not $i$-defend $\phi$, then $(i, \phi, \hist, \ar_\gamma)$ falsifies $\gamma$, which ends the proof. 
	Thus, assume that $\ar_\gamma$ $i$-defends $\phi$.

@@

\begin{theorem}
	Given two incompatible empirical theories for a given normative theory, $\alpha = (\gleadsto[\alpha], \gbeats[\alpha], \alldargs, \dbeats[\alpha], +_\alpha)$ and $\beta = (\gleadsto[\beta], \gbeats[\beta], \alldargs, \dbeats[\beta], +')$ and given any $i \in I$, there exists a test procedure with $i$ that falsifies one of these theories.
\end{theorem}
\begin{proof}
	Pick any $\phi \in \Phi$ about which $\alpha$ and $\beta$ disagree.

@@

	Pick any $\ar_{-1} \in \allargs \suchthat \ar_{-1} \gleadsto[\beta] ¬\phi$. Observe that $\ar_{-1} \in \gargsbeta \cap (\dfp[\beta])^0(\gleadstoinv[\beta](¬\phi))$. Test whether $(\ar_{-1}) \nileadstost \phi$. If so, $(i, \phi, (\ar_{-1}), \ar_{-1})$ falsifies $\beta$, which ends the procedure. 
	Thus, assume that $(\ar_{-1}) \ileadstost \phi$.

	Pick any $\ar_0 \in \allargs \suchthat \ar_0 \gleadsto[\alpha] \phi$. Observe that $\ar_0 \in \gargsalpha \cap (\dfp[\alpha])^0(\gleadstoinv[\alpha](\phi))$.  Define $\hist = (\ar_{-1}, \ar_0)$ and test whether $\hist \nileadstost \phi$. By \cref{th:endchal}, $(\hist, \ar_0, \ar_{-1})$ is a $(\alpha, \phi$ or $\beta, ¬\phi)$-challenge, thus, $(\hist, \ar_0, \ar_{-1})$ is a challenge of complexity $(0, 0)$.

@@

Define $\ileadstost$ as follows. Given $\hist = (…, \ar)$, $\hist \ileadstost \phi$ iff, in a situation of $i$ having been exposed to the prefix …, saying: “Considering the argument $\ar$, would you please indicate your maximum selling price for $P$, and for $D$” leads $i$ to act accordig to $\phi$. Note that $\hist \ileadstost \phi$ iff $\hist \nileadstost ¬\phi$.
